---
title: "Tutorial - pseqML"
output: html_document
date: "2023-02-07"
---

## Before to start

```{r}
# You need to have installed the following packages
require(mlr3)
require(mlr3learners)
require(mlr3verse)
require(caret)
require(phyloseq)

# Then, you can load pseqML
require(pseqML)
```

## Import ```phyloseq``` object
Before to start working with ```pseqMLR```, you need to create a ````phyloseq``` object. See the oficial [tutorial](https://joey711.github.io/phyloseq/) for more information. You can use our toy dataset to follow this tutorial.

```{r}
data("random_pseq")
print(pseq)
```
In this toy example (generated randomly), we have a ```phyloseq``` object that includes eight taxons through 20 samples. In addition, samples are labeled according two variables.

```{r}
names(sample_data(pseq))
```


## Preprocessing steps
The first step that you need to do is convert ```phyloseq``` object into a ```pseqMLR``` object.

```{r}
pseqML <- pseq2mlr(
  phyloseq = pseq,
  target = "cancer",
  id = "example")
```

Normally, When we are working with a ```phyloseq``` object, we need to make several transformations and preprocessing steps before to build our prediction model. ```pseqML``` facilitates this task by unifying in a single package, the state-of-the-art preprocessing methods in an easy and intuitite interface, working mainly on the ```pseqMLR``` object. Below are the functions included in ```pseqML```. These functions can be divided into XX groups:

1. Functions to transform counts
2. Functions to filter taxonomic features
3. Functions to filter samples
4. Functions to make descriptive analysis
5. Functions to plot

### Functions to transform counts
```pseqMLR``` works ideally with raw counts. This raw counts can be transformed through multiple ways. 

```{r}
# Transform counts
relab <- relative_abundance(pseqML)
otu_table(relab@data)

zscore <- zscore_abundance(pseqML)
otu_table(zscore@data)

deep <- deep_median_abundance(pseqML)
otu_table(deep@data)

t <- transform_counts(pseqML, "clr")
otu_table(t@data)
```

### Functions to filter taxonomic features
```{r}
# Filter taxonomic features
phyloseq::ntaxa(pseqML@data)
pseqML_f <- filtering_taxa(
  pseqMLR = pseqML,
  type = "presence",
  threshold = 1,
  percentage = 0.5)
phyloseq::ntaxa(pseqML_f@data)
```


### Functions to filter samples
```{r}
# Filter samples
phyloseq::nsamples(pseqML@data)
pseqML_fs <- filtering_counts(pseqML, 12)
phyloseq::nsamples(pseqML_fs@data)
```


### Functions to make descriptive analysis
```{r}
# Checking steps
check_prevalence(pseqML, level = "Rank6")
```


### Functions to plot
```{r}
plot_abundances(pseqML, level = "Rank5")              # Warning!
plot_abundances_target(pseqML, level = "Rank5")       # Warning!
plot_richness(pseqML, index = "shannon")
plot_abundances_bars(pseqML, level = "Rank6", transformation = "none") # warning
plot_abundances_lines(pseqML, level = "Rank6") # warining!
plot_heatmap(pseqML, level = "Rank6", transformation = "log") # warning!
plot_varcorr(pseqML, target = NULL, level = "Rank6", method = "pearson")  #error!
plot_taxa_core(pseqML, target = NULL)  # warning!


```

